---
title: "ds202_lab3_team"
author: 'Jessie Bustin Github: jebustin'
date: "2/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
df <- readxl::read_xls("GSS.xls")
install.packages("dplyr")
install.packages("ggplot2")
library("dplyr")
df
```

Data Cleaning:
1.To being with with cleaning I used "dplyr" in R. 
2.ColNames changed for easier referencing. 
3.Dropped Na values for num col (Year and ID only)
4.Dropped any rows containing "No answer" or "Don't know" or "Dk na".
5.I didn't remove any "Not applicable" rows
```{r}
df1 = rename(df, "Year"='Gss year for this respondent',
             "Pol_Affl"="Political party affiliation",
             "Race"="Race of respondent",
             "Sex"="Respondents sex",
             "respondentSpouse_TopEdu"="Spouses highest degree",
             "Respondent_TopEdu"="Rs highest degree",
             "YearHSCompleted"="Highest year of school completed",
             "NumOfChildren"="Number of children",
             "WidowStatus"="Ever been widowed",
             "MaritalStatus"="Marital status", 
             "HoursWorked"="Number of hours usually work a week",
             "ID"="Respondent id number",
             "HappyIndex"="General happiness")
df1<-na.omit(df1)
df1<-df1[!(df1$Pol_Affl=="No answer" | df1$respondentSpouse_TopEdu=="No answer" | df1$respondentSpouse_TopEdu=="Don't know" | df1$Respondent_TopEdu=="No answer" | df1$YearHSCompleted=="No answer" | df1$NumOfChildren=="Dk na" | df1$WidowStatus=="No answer" | df1$WidowStatus=="Don't know" | df1$MaritalStatus=="No answer" | df1$HoursWorked=="No answer" | df1$HoursWorked=="Don't know" | df1$HappyIndex=="No answer" | df1$HappyIndex=="Not applicable"),]
df1
```

## 1

How does the happiness of a respondent relate to the marriage status?
The bar plot below shows that people are genrally more happier while they're married. So are people who've ne'er married. But they're people who only are pretty happy not VERY happy, being divorced.

```{r}
library("ggplot2")
  Q1<-df1 %>%
  select(MaritalStatus, HappyIndex)
Q1
ggplot(data=Q1,aes(x=MaritalStatus,fill=HappyIndex))+geom_bar()
```


## 2

Does the sex of the respondent affect the relationship you found in Q1?

```{r}
 Q2<-df1 %>%
  select(MaritalStatus, HappyIndex, Sex)
```

## 3

How does the happiness of a respondent relate to the political party affiliation?

```{r}

```

## 4

Raise four questions that can be answered by the GSS dataset. Extract appropriate data from the website, explore, and answer the questions.

1. Does the number of children the respondent has vary with race?
Does this relationship change over the years?

```{r}

```

2. Does the respondents spouse's highest degree affect the years of
school the respondent completed? Does this relationship change for
for men and women?

```{r}

```

3. Does the distribution between political parties change based on
the highest degree of the respondent?

```{r}

```

4. Is there a relationship between happiness and number of hours
usually worked in a week?  Does this conclusion hold true for both
men and women?

```{r}

```
